Unsupervised Semantic Mapping for Healthcare Data Storage Schema

### 2021-06-04 ManuScript/IEEEAccess/ManuScript.tex
update discussion 1

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

@@

\begin{abstract}
Data, information, and knowledge processing systems, in the domain of healthcare are currently plagued by the heterogeneity at various levels. Current solutions, have focused on developing a standard based, manual intervention mechanism, which requires a large amount of human resource and necessitate realignment of existing systems. State-of-the-art methodologies in the field of natural language processing and machine learning, can help to partially automate this process, reducing the resource requirements and providing a relatively good multi-class based classification algorithm. We present a novel methodology for bridging the gap between various healthcare data management solutions by leveraging the strength of transformer based machine learning models, to create mappings between the data elements. Additionally, the annotated data, collected against five medical schemas and labeled by 4 annotators is made available for helping future researchers. Our results indicate, that for biased, dependent multi-class text classification, RoBERTa shows the best performance, achieving a cohen's kappa score of 0.47 and MCC score of 0.48, with human annotated data.
\end{abstract}

\begin{keywords}
Electronic medical records, Health information management, Medical information systems,  Text processing, Unsupervised learning
\end{keywords}

@@

Data interoperability resolves the heterogeneity between data artificats, to enable, seamless and interpretable communcation among source and target organizations, while preserving the data's original intention during storage, communication, and usage (as defined by IEEE 610.12 \cite{geraci1991ieee}, Health Level Seven International (HL7), and Healthcare Information and Management Systems Society HIMSS \cite{Himss_url2013}).

@@

Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards, necessitates the creation of bridging standards, that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (CIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario, requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 

@@

A similar notion of data heterogeneity, in terms of medical data schema, is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name" and ``patientName", which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others, such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address", ``City", ``State", ``Postal Code", ``Country", and others. Similarly, ``EncounterDate" from $s_5$ is semantically similar to ``BeginDate" of ``openemr\_MedicalProblems" table in $s_1$, ``AdmissionStartDate" of ``LabsCorePopulatedTable" in $s_2$, 
``time" in ``Diagnosis" table of $s_3$, and  ``dateOfAdmission" in ``Diagnosis" and ``CareHistory" tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

@@

\subsection{From Attribute to Sentence}
\label{schema_prepro}
In order to process the EMR/EHR schema set $S$ and produce a set of corresponding semantically enriched sentences, we use the data representation $s_i$, generated through the process explained in sequence acquisition to collect the various medical fragments in memory. We then iterate over these fragments, building a set of attributes, distinguished by their name, schema's name, table's name, schema's version, source, and recorded data. This entails that ``PatientID" from each of the four tables in $s_2$, and ``patientID" from five tables in $s_4$, would result into nine attributes (assuming, as in the current case, of no differences in versions of these systems).
For each attribute, we then generate the suffix array, which provides all possible substring representations contained within the attribute name. In order to generate the set of suffixes, we employ three strategies, forward suffix generation, whereby for a word $w$ of length $n$, $n-1$ suffixes of size 2 to $n-1$ are produced, backward suffix generation, to produce $n-1$ suffixes in reverse order with size $n-1$ to 2, and regular expression based suffix generation, which splits each word on, change of case, special characters(such as -, \_, !, and others), and numbers. In this way a large list of suffixes is generated, which is combined using a ``TreeSet" data structure of Java, which internally sorts this list as well. An example of this suffix generation process, using the attribute name ``dateOfAdmission" as it appears in $s_4$ is shown in \ref{fig:suffix_attribute}.

Suffix strings for similar attributes such as ``AdmissionStartDate". ``diseaseNameOnAdmission", and ``AdmissionEndDate" appear in $s_2$, produce many, synctactically similar suffixes, to the presented example. This process, is only able to generate syntactic suffixes, producing many incoherent and unrelated suffixes. In order to counter this problem, and to limit the list of suffixes within the domain, we then query UMLS, with exact search strategy, looking for the existence of any concepts, against each suffix. In case, no semantic concept is found for a particular suffix, it is removed from the final Suffix Array. On the other hand, if atleast one semantic concept is found against the queried suffix, it is retained. Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. 
The set of suffixes and their corresponding concepts are then used to build the sentence, where by each concept, corresponding to a suffix is appended next to the suffix. An example of the resultant sentence for the attribute ``DateOfAdmission" is shown as follows:\\

\framebox{\parbox{\dimexpr\linewidth-2\fboxsep-2\fboxrule}{\itshape%

		``Date Value type - Date date allergenic extract Date in time Data types - Date Date Fruit;Of SPI1 wt Allele SPI1 gene TAF1 wt Allele BRIP1 gene Within Degrees fahrenheit Oral contraception BRIP1 wt Allele;Da Displacement of abomasum dalton Anterior descending branch of left coronary artery deca units cytarabine/daunorubicin protocol Dai Chinese Asymptomatic diagnosis of Drug Accountability Domain;ion Iontophoresis Route of Drug Administration Ions;on SPARC wt Allele Osteonectin SPARC gene On (qualifier value) Upon - dosing instruction fragment;Admission Admission activity Hospital admission;Dat SLC6A3 gene SLC6A3 wt Allele dopamine transporter Direct Coombs test SLC6A3 protein, human Test Date cytarabine/daunorubicin/thioguanine Alzheimer's Disease;mission Religious Missions;"
}}\\

Here the various suffixes and their concepts are seperated by the symbol ``;", however together they form one sentence, for which an embedded vector is generated.

@@

Schema Maps, provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences, corresponding to each data attribute. While the embedded vectors can be generated using any methodology (we tested 11 methodologies, with WordNet and 10 models based on BERT (including BERT base, BERT large, distilBert, and RoBERTa) further detailed in section \ref{experimentalSetup}), the large/STSb version of RoBERTa\cite{reimers-2019-sentence-bert},provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates, inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal" (the two attributes are same), ``related" (the two attributes are related to each other), and ``unrelated" (no relationship between the attributes). In order to classify the similarity results, into one of these three classes, we then calculated the best thresholds, using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal", ``related", and ``unrelated".

\begin{equation}
MCC = \frac{(TP.FN) - (FP.FN)}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\longrightarrow [-1,1]
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives, MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the mcc score towards -1. This measurement is markedly different from accuracy, that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate, in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated") to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach, using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments, follow various design elements, supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation"\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files, containing an index for patients, an index for their medical fragments, and the medical fragment, corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a ``json" file, which is then read by the same application to partially generate the schema maps. This process, is used to create 20,349 distinct pairs of attributes, across $s$. Each pair also contains the ``relationshipList", which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The ``json" file thus produced, is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include 'bert-base-nli-stsb-mean-tokens', 'bert-large-nli-stsb-mean-tokens', 'roberta-base-nli-stsb-mean-tokens', 'roberta-large-nli-stsb-mean-tokens', 'distilbert-base-nli-stsb-mean-tokens', 'bert-base-nli-mean-tokens', 'bert-large-nli-mean-tokens', 'roberta-base-nli-mean-tokens', 'roberta-large-nli-mean-tokens', and 'distilbert-base-nli-mean-tokens'. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages, is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools. These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth, and to identify the best thresholds for classifying each instance as ``equal", ``related", or ``unrelated" 4 human annotators were utilized, to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix, by marking all attribute pairs belonging to the same schema with the symbol ``-". Following this, the annotators, marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells, with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0", exactly similar as ``1", row attribute as child of column attribute as ``<", row attribute as a parent of the column attribute as ``>", and finally, unknown as ``~". The data sheets generated after this extensive human effort, have been made available for other researchers\footnote{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}.

@@

These sheets, additionally contain some missing values, which were left out by the annotators, but in order to maintain their originality, these values were not filled; instead during our evaluation for these datasets, the missing values were considered as having the score ``0". Using $\kappa$, we evaluated the inter-rater agreement of these annotations, which have been visualized in Figure \ref{fig:kappaInterAnnotatorAgrement}. It can be seen in this plot, that ``Annotator3" has very small correlation with the other 3 annotators. This difference can be traced back to the number and type of annotations performed by each annotator, which is shown in Table \ref{tab:annotator_marking}. The ``Annotator3" has marked 2103 cells as related (one of >, < , or ~) and left 153 as empty. However, even in the presence of these difference, it is pertinent to include the data for all annotators in order to avoid any biasedness.

This annotated data was then processed, to replace all related entries with ``0.5" ($class_{0.5}$) and all``-" with ``0" ($class_{0}$), while the values for similar at ``1" ($class_{1}$) and ``0" ($class_{0}$) for dissimilar were kept the same. This conversion was then used to produce a consolidated dataset of 40,698 attribute pairs, using mode scores of all annotators, for each cell. We also tested average scores between the annotators, but that would produce scores between ``0", ``0.5", and ``1", greatly increasing the number of classes for classification. Hence the maximum agreement between the annotators, maintains the final label values within these three classes, which become easier to evaluate. Additionally, the original dataset and its mode consolidated form is biased in favour of class "0", since most attribute pairs are not related to each other.
This dataset is then split into development and testing partitions with a ratio of 70:30. The development partition is used for threshold selection, based on the best MCC score for identifying class ``equal", followed by best scores for class ``related" and finally best of class "unrelated". The optimal threshold thus achieved is used to classify the instances of the test dataset, which is finally evaluated on its MCC and $F_1$ measure.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the Matthews Correlation Coefficient (MCC), is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem ($class_0$, $class_{0.5}$, and $class_1$), we have to test various threshold scores for separating the instances between $class_0$ and $class_{0.5}$ ($t_1$), and then $class_{0.5}$ and $class_1$ ($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of $class_1$ (similar), followed by $class_{0.5}$ (related), and finally $class_0$ (unrelated).
With a step size of \textit{0.05} ($step$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step$, we move the thresholds until $t_2$ reaches \textit{1.0}, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches \textit{0.95} and $t_2$ reaches \textit{1.0}, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy", ``Word2Vec", and 10 BERT based models.

@@

		& $class_{0.5}$                     & 0.0,$class_1$                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 

@@

		& $class_{0.5}$                     & 0.0,1                   & 1.00                                   & 0.00                                    & 0.00                                 & 0.00                              & 0.00                              \\ \cline{2-8} 

>>>>

\documentclass{ieeeaccess}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{ulem}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\history{Date of publication xxxx 00, 0000, date of current version xxxx 00, 0000.}
\doi{10.1109/ACCESS.2017.DOI}

@@

\begin{abstract}
Data, information, and knowledge processing systems, in the domain of healthcare are currently plagued by the heterogeneity at various levels. Current solutions, have focused on developing a standard based, manual intervention mechanism, which requires a large amount of human resource and necessitate realignment of existing systems. State-of-the-art methodologies in the field of natural language processing and machine learning, can help to partially automate this process, reducing the resource requirements and providing a relatively good multi-class based classification algorithm. We present a novel methodology for bridging the gap between various healthcare data management solutions by leveraging the strength of transformer based machine learning models, to create mappings between the data elements. Additionally, the annotated data, collected against five medical schemas and labeled by 4 annotators is made available for helping future researchers. Our results indicate, that for biased, dependent multi-class text classification, RoBERTa shows the best performance, achieving a cohen's kappa score of 0.47 and Matthews Correlation Coefficient (MCC) score of 0.48, with human annotated data.
\end{abstract}

\begin{keywords}
\sout{Electronic medical records, Health information management, Medical information systems,  Text processing, Unsupervised learning}
\end{keywords}

@@

Data interoperability resolves the heterogeneity between data artificats to enable seamless and interpretable communcation among source and target organizations, while preserving the data's original intention during storage, communication, and usage (as defined by IEEE 610.12 \cite{geraci1991ieee}, Health Level Seven International (HL7), and Healthcare Information and Management Systems Society HIMSS \cite{Himss_url2013}).

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards, necessitates the creation of bridging standards, that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (CIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario, requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 

@@

A similar notion of data heterogeneity, in terms of medical data schema, is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name'' and ``patientName'', which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others, such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address'', ``City'', ``State'', ``Postal Code'', ``Country'', and others. Similarly, ``EncounterDate'' from $s_5$ is semantically similar to ``BeginDate'' of ``openemr\_MedicalProblems'' table in $s_1$, ``AdmissionStartDate'' of ``LabsCorePopulatedTable'' in $s_2$, 
``time'' in ``Diagnosis'' table of $s_3$, and  ``dateOfAdmission'' in ``Diagnosis'' and ``CareHistory'' tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

@@

\subsection{From Attribute to Sentence}
\label{schema_prepro}
In order to process the EMR/EHR schema set $S$ and produce a set of corresponding semantically enriched sentences, we use the data representation $s_i$, generated through the process explained in sequence acquisition to collect the various medical fragments in memory. We then iterate over these fragments, building a set of attributes, distinguished by their name, schema's name, table's name, schema's version, source, and recorded data. This entails that ``PatientID'' from each of the four tables in $s_2$, and ``patientID'' from five tables in $s_4$, would result into nine attributes (assuming, as in the current case, of no differences in versions of these systems).
For each attribute, we then generate the suffix array, which provides all possible substring representations contained within the attribute name. In order to generate the set of suffixes, we employ three strategies, forward suffix generation, whereby for a word $w$ of length $n$, $n-1$ suffixes of size 2 to $n-1$ are produced, backward suffix generation, to produce $n-1$ suffixes in reverse order with size $n-1$ to 2, and regular expression based suffix generation, which splits each word on, change of case, special characters(such as -, \_, !, and others), and numbers. In this way a large list of suffixes is generated, which is combined using a ``TreeSet'' data structure of Java, which internally sorts this list as well. An example of this suffix generation process, using the attribute name ``dateOfAdmission'' as it appears in $s_4$ is shown in \ref{fig:suffix_attribute}.

Suffix strings for similar attributes such as ``AdmissionStartDate''. ``diseaseNameOnAdmission'', and ``AdmissionEndDate'' appear in $s_2$, produce many, synctactically similar suffixes, to the presented example. This process, is only able to generate syntactic suffixes, producing many incoherent and unrelated suffixes. In order to counter this problem, and to limit the list of suffixes within the domain, we then query UMLS, with exact search strategy, looking for the existence of any concepts, against each suffix. In case, no semantic concept is found for a particular suffix, it is removed from the final Suffix Array. On the other hand, if atleast one semantic concept is found against the queried suffix, it is retained. Meanwhile the process continues for the next attribute, then the next table, and finally the next system, till no further processing is possible. 
The set of suffixes and their corresponding concepts are then used to build the sentence, where by each concept, corresponding to a suffix is appended next to the suffix. An example of the resultant sentence for the attribute ``DateOfAdmission'' is shown in Table \ref{tab:sentence_example}.

\begin{table}[]
	\centering
	\caption{Sentence created from the attribute name ``DateOfAdmission''}
	\label{tab:sentence_example}
	\begin{tabular}{p{\linewidth}}
		Date Value type - Date date allergenic extract Date in time Data types - Date Date Fruit;Of SPI1 wt Allele SPI1 gene TAF1 wt Allele BRIP1 gene Within Degrees fahrenheit Oral contraception BRIP1 wt Allele;Da Displacement of abomasum dalton Anterior descending branch of left coronary artery deca units cytarabine/daunorubicin protocol Dai Chinese Asymptomatic diagnosis of Drug Accountability Domain;ion Iontophoresis Route of Drug Administration Ions;on SPARC wt Allele Osteonectin SPARC gene On (qualifier value) Upon - dosing instruction fragment;Admission Admission activity Hospital admission;Dat SLC6A3 gene SLC6A3 wt Allele dopamine transporter Direct Coombs test SLC6A3 protein, human Test Date cytarabine/daunorubicin/thioguanine Alzheimer's Disease;mission Religious Missions;
	\end{tabular}
\end{table}

Here the various suffixes and their concepts are separated by the symbol ``;'', however together they form one sentence, for which an embedded vector is generated.

@@

Schema Maps, provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences, corresponding to each data attribute. While the embedded vectors can be generated using any methodology (we tested 11 methodologies, with WordNet and 10 models based on BERT (including BERT base, BERT large, distilBert, and RoBERTa) further detailed in section \ref{experimentalSetup}), the large/STSb version of RoBERTa\cite{reimers-2019-sentence-bert},provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates, inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal'' (the two attributes are same), ``related'' (the two attributes are related to each other), and ``unrelated'' (no relationship between the attributes). In order to classify the similarity results, into one of these three classes, we then calculated the best thresholds, using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal'', ``related'', and ``unrelated''.

\begin{equation}
\resizebox{0.9\hsize}{!}{$%
	MCC = \frac{(TP.FN) - (FP.FN)}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\longrightarrow [-1,1]
	$%
}%
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives, MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the mcc score towards -1. This measurement is markedly different from accuracy, that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate, in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated'') to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach, using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments, follow various design elements, supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation''\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files, containing an index for patients, an index for their medical fragments, and the medical fragment, corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a JSON file, which is then read by the same application to partially generate the schema maps. This process, is used to create 20,349 distinct pairs of attributes, across $s$. Each pair also contains the ``relationshipList'', which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The JSON file thus produced, is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include ``bert-base-nli-stsb-mean-tokens'', ``bert-large-nli-stsb-mean-tokens'', ``roberta-base-nli-stsb-mean-tokens'', ``roberta-large-nli-stsb-mean-tokens'', ``distilbert-base-nli-stsb-mean-tokens'', ``bert-base-nli-mean-tokens'', ``bert-large-nli-mean-tokens'', ``roberta-base-nli-mean-tokens'', ``roberta-large-nli-mean-tokens'', and ``distilbert-base-nli-mean-tokens''. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages, is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools.
%These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.%

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth, and to identify the best thresholds for classifying each instance as ``equal'', ``related'', or ``unrelated'' 4 human annotators were utilized, to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix, by marking all attribute pairs belonging to the same schema with the symbol ``-''. Following this, the annotators, marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells, with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0'', exactly similar as ``1'', row attribute as child of column attribute as ``<'', row attribute as a parent of the column attribute as ``>'', and finally, unknown as ``~''. The data sheets generated after this extensive human effort, have been made available for other researchers\footnote{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}.

@@

These sheets, additionally contain some missing values, which were left out by the annotators, but in order to maintain their originality, these values were not filled; instead during our evaluation for these datasets, the missing values were considered as having the score ``0''. Using $\kappa$, we evaluated the inter-rater agreement of these annotations, which have been visualized in Figure \ref{fig:kappaInterAnnotatorAgrement}. It can be seen in this plot, that ``Annotator3'' has very small correlation with the other 3 annotators. This difference can be traced back to the number and type of annotations performed by each annotator, which is shown in Table \ref{tab:annotator_marking}. The ``Annotator3'' has marked 2103 cells as related (one of >, < , or ~) and left 153 as empty. However, even in the presence of these difference, it is pertinent to include the data for all annotators in order to avoid any biasedness.

This annotated data was then processed, to replace all related entries with ``0.5'' ($class_{0.5}$) and all``-'' with ``0'' ($class_{0}$), while the values for similar at ``1'' ($class_{1}$) and ``0'' ($class_{0}$) for dissimilar were kept the same. This conversion was then used to produce a consolidated dataset of 40,698 attribute pairs, using mode scores of all annotators, for each cell. We also tested average scores between the annotators, but that would produce scores between ``0'', ``0.5'', and ``1'', greatly increasing the number of classes for classification. Hence the maximum agreement between the annotators, maintains the final label values within these three classes, which become easier to evaluate. Additionally, the original dataset and its mode consolidated form is biased in favour of class ''0'', since most attribute pairs are not related to each other.
This dataset is then split into development and testing partitions with a ratio of 70:30. The development partition is used for threshold selection, based on the best MCC score for identifying class ``equal'', followed by best scores for class ``related'' and finally best of class ''unrelated''. The optimal threshold thus achieved is used to classify the instances of the test dataset, which is finally evaluated on its MCC and $F_1$ measure.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the Matthews Correlation Coefficient (MCC), is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem ($class_0$, $class_{0.5}$, and $class_1$), we have to test various threshold scores for separating the instances between $class_0$ and $class_{0.5}$ ($t_1$), and then $class_{0.5}$ and $class_1$ ($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of $class_1$ (similar), followed by $class_{0.5}$ (related), and finally $class_0$ (unrelated).
With a step size of \textit{0.05} ($step$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step$, we move the thresholds until $t_2$ reaches \textit{1.0}, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches \textit{0.95} and $t_2$ reaches \textit{1.0}, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy'', ``Word2Vec'', and 10 BERT based models.

@@

@@
